---
title: Supplement
output:
  bookdown::pdf_document2:
    extra_dependencies: ["colortbl"]
    toc: false
    number_sections: false
  bookdown::html_document2:
      toc: true
      number_sections: false
  bookdown::word_document2: default
always_allow_html: yes
---

```{r si-set-up, include=FALSE}
# set up -----------------------------------
# Packages
library(knitr)
library(here)
library(dplyr)
library(ggplot2)
library(kableExtra)

# Document settings
opts_chunk$set(eval = TRUE, echo = FALSE,
               message = FALSE, warning = FALSE,
               eval.after = "fig.cap")
options(scipen=1, digits=2)
theme_set(theme_bw())

# Set latest date for evaluation
eval_date <- as.Date("2022-03-07")

# Load data: evaluation scores
load_from_local <- TRUE # load from saved csv in "data" (FALSE starts download)
```

Supplementary information: _Predictive performance of multi-model ensemble forecasts of COVID-19 across European nations_

```{r get-si, include = FALSE}
# Get SI figures
source(here("code", "summarise", "supplementary.R"))
```

### Participating teams

The following teams participated in the European Forecast Hub by contributing forecasts over the study period. Information below is taken from metadata provided by each team.

```{r team-table}
team_table
# team_table %>%
#   kable(format = "markdown", booktabs = T,
#         caption = "Teams participating in the European Forecast Hub contributing a model between March 2021 - March 2022") %>%
#   kableExtra::kable_styling(kable_input = .,
#                             full_width = TRUE,
#                             latex_options = c("striped", "scale_down"))
```

 \newpage
 
### Summary of evaluated forecasts
```{r si-fig-1, fig.cap=si_fig1_legend, fig.height=6, fig.width=5}
si_figure_1
```

\newpage
 
### Comparison of contributed forecasts and the Hub ensemble
```{r si-fig-2, fig.cap=si_fig2_legend, fig.height=3, fig.width=7}
si_figure_2
```

